---
title: "Tidy Data"
output: github_document
---

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

(what about 'gather()'..? never use this one. Always use 'pivot_longer()')

## pivot longer

Load the PULSE data

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Let's try to pivot

```{r}
pulse_tidy = 
  pulse_df %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_", 
    values_to = "bdi"
  ) %>% 
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)
  )
```

## pivot_wider

make up a results data table
```{r}
analysis_df =
  tibble(
    group = c("treatment", "treatment", "control", "control"),
    time = c("a", "b", "a", "b"),
    group_mean = c(4, 8, 3, 6)
  )

analysis_df %>% 
  pivot_wider(
    names_from = "time",
    values_from = "group_mean"
  ) %>% 
knitr::kable()
```

(what about 'spread'..? never use this one. Always use 'pivot_wider')


#learning assessment 1

```{r}
litters_wide = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(litter_number, ends_with("weight")) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd", 
    values_to = "weight") %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))

```

## bind_rows

Import the LotR movie words stuff

```{r}
fellowship_df =
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_rings")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")

lotr_tidy = 
  bind_rows(fellowship_df, two_towers, return_king) %>% 
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words"
  ) %>% 
  relocate(movie)
```

(what about 'rbind'..? never use this one. Always use 'bind_rows')


## joins

Loo at FAS data. This imports and cleans litters and pups data. 
```{r}
litter_df =
  read_csv("data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>% 
  relocate(litter_number) %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))

pups_df = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, '1' = "male", '2' = "female"))
```
  
Let's joint these up

```{r}
fas_df = 
  left_join(pups_df, litter_df, by = "litter_number") %>% 
  relocate(litter_number, dose, day_of_tx)
```
















